
\documentclass[10pt,letterpaper, notitlepage]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[noabbrev]{cleveref}
\usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{authblk}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\usepackage{todonotes}

%%
\linespread{1.5}
\newcommand{\Normal}{\operatorname{Normal}}
\newcommand{\Gmma}{\operatorname{Gamma}}
\newcommand{\Lognormal}{\operatorname{Lognormal}}

%Import the natbib package and sets a bibliography  and citation styles
\usepackage[round, sort, numbers]{natbib}
\setcitestyle{square}
\bibliographystyle{unsrtnat}
%%

\title{Developing and Evaluating Pharmacokinetics-Driven \\Dynamic Personalized Medicine: A Framework and Case Study}

%\author{A. Demetri Pananos M.Math, Rommel G. Tirona Ph.D, Simon Bonner Ph.D,  Daniel J. Lizotte Ph.D}
\author[1, *]{A. Demetri Pananos, M.Math}
\author[2]{Rommel G. Tirona Ph.D}
\author[3]{Simon Bonner Ph.D}
\author[1,4]{Daniel J, Lizotte, Ph.D}
\affil[1]{Department of Epidemiology and Biostatsitics, Western University}
\affil[2]{Department of Physiology and Pharmacology, Western University}
\affil[3]{Department of Statistics and Actuarial Sciences, Western University}
\affil[4]{Department of Computer Science, Western University}
\affil[*]{Corresponding Author}

\setcounter{Maxaffil}{0}
\date{}


\begin{document}
	\begin{titlepage}
	\maketitle
	
	Word count: 4496
	\footnote{Financial support for this study was provided by Ontario Research Fund and Natural Sciences and Engineering Research Council of Canada. The funding agreement ensured the authors’ independence in designing the study, interpreting the data, writing, and publishing the report.}
	\end{titlepage}

	
	% Second title page
	\begin{titlepage}
		\centering
		\vskip 60pt
		\topskip0pt
		\vspace*{\fill}
		\LARGE Developing and Evaluating Pharmacokinetics-Driven \\Dynamic Personalized Medicine: A Framework and Case Study \par
		\vskip 3em
		\vspace*{\fill}
		\large %Author A, Author B \par
		%\vskip 1.5em
		Acknowledgements: The authors wish to thank A.J.B for their comments on health economic impacts of this work, as well as R.K and their lab for providing pharmacological expertise.
	\end{titlepage}

\begin{abstract}
\textbf{Background}: One goal of personalized medicine is to select the right dose for a drug by considering factors which may explain variability in effectiveness or toxicity. Most efforts to personalize dose size seek to inform dose selection at a single time, with the goal of eliminating the need for dose adjustments (titration). We refer to this as static personalization. Despite progress, the need for titration has not been eliminated and thus there is an opportunity to personalize the titration process as well. We refer to this as dynamic personalization. Despite its potential to improve care, dynamic personalization imposes additional burden on the patient and provider, because it requires ongoing monitoring, for example by gathering lab results and requiring additional clinic visits. It is natural to ask:  is the additional control over dose worth the additional burden?\textbf{Objective}: To develop a framework for the simulation based evaluation of static and dynamic personalization based on pharmacokinetic modelling.Methods: We frame dynamic personalization as a dynamic treatment regime and describe how Bayesian pharmacokinetic models can be used to estimate the optimal policy with Q learning, and we propose a simulation-based strategy for internal evaluation. \textbf{Results}: We demonstrate the use of our framework to investigate the potential benefits of static and dynamic personalization for apixaban dosing. \textbf{Conclusions}: The proposed framework can help compare potential burden and benefits of different modes of personalized dosing and it can be integrated with other system level decision making frameworks in order to evaluate whether such a personalized medicine program should be implemented into a particular health care system.


\end{abstract}

	\input{sections_slim/001_introduction.tex}
	\input{sections_slim/002_background.tex}
	\input{sections_slim/003_framework.tex}
	\input{sections_slim/004_results.tex}
	\input{sections_slim/005_discussion_and_conclusion.tex}
	\newpage
	\bibliography{./bibs/introduction,./bibs/dtr, ./bibs/bayes}
	\newpage
	\appendix
	\input{sections_slim/00A_appendix.tex}

\end{document}